BIP-Predictor

### 2025-05-07 Latex/BIP_predictor.tex
up

Extending the `classic' direct forecast design, we here consider filtered indicators, based on the one-sided HP(160), denoted as HP-C: the filtered indicators are explanatory variables for the regressions on forward-shifted BIP\footnote{Katja and Simon: for each forecast horizon $h=0,1,2,3$ I'm using corresponding forecasts of the real-time filter output, assuming the data to be white noise (assuming an ARMA model does not lead to significantly different results). I think we might ignore these technical details in the paper.}. As in the previous section, we rely on all indicators as well as the full sample estimate for illustration, see Figs.\eqref{direct_hp_forecasts} (all observations with exclusion of the pandemic) and \eqref{direct_hp_forecasts_financial_crisis} (financial crisis). A comparison with the classic direct forecasts in Fig.\eqref{direct_wc_financial_crisis} indicates that the new predictors are slightly less retarded (`faster'). 
\begin{figure}[H]\begin{center}\includegraphics[height=3in, width=4.5in]{./Figures/direct_hp_forecasts.pdf}\caption{Direct HP forecasts: entire data set with exclusion of the pandemic.\label{direct_hp_forecasts}}\end{center}\end{figure}\begin{figure}[H]\begin{center}\includegraphics[height=3in, width=4.5in]{./Figures/direct_hp_forecasts_financial_crisis.pdf}\caption{Direct HP forecasts over the curse of the financial crisis.\label{direct_hp_forecasts_financial_crisis}}\end{center}\end{figure}Moreover, the F-statistics are larger than for the direct forecasts, see Table \eqref{f_stat}, but the predictor remains statistically insignificant at forecast horizons exceeding two quarters (even though the evaluation is conducted in-sample and based on all available indicators). 
% latex table generated in R 4.2.2 by xtable 1.8-4 package
% Thu May  1 21:38:41 2025
\begin{table}[ht]
\centering
\begin{tabular}{rrrrr}
  \hline
 & h= 0 & h= 1 & h= 2 & h= 3 \\ 
  \hline
Direct forecast & 7.188 & 5.285 & 3.574 & 0.460 \\ 
  Direct HP forecast & 21.886 & 8.415 & 4.810 & 1.834 \\ 
   \hline
\end{tabular}
\caption{F-statistics of direct forecasts and direct HP forecasts for forecast horizons $h=0,1,2,3$: fully in-sample.  } 
\label{f_stat}
\end{table}We now propose a more refined causal filter design aiming for statistical significance of the resulting predictor at longer forecast horizons, out-of-sample.

@@

% Thu May  1 21:38:41 2025

@@

Target correlation & 0.55 & 0.63 & 0.59 & 0.68 \\ 
  HT & 8.67 & 6.50 & 8.67 & 4.00 \\ 

@@

M-SSA predictors are compared to the traditional univariate HP-C in Fig. \eqref{multivar_vs_univar} across the periods of the financial crisis (top panels) and the COVID-19 pandemic (bottom panels), for forecast horizons $h=0,...,6$. The nowcast from the previous section is included for reference. All series are standardized to facilitate visual comparison. The primary distinction of the M-SSA approach is the progressive left-shift of the predictor as a function of $h$; this shift becomes more pronounced with increasing $h$ and appears to be consistent across all levels, including the peaks and troughs of the series. Conversely, the positions of peaks and troughs in the HP-C series appear to remain largely unaffected by variations in $h$. The more pronounced and systematic left-shift observed in the M-SSA forecasts can be partly attributed to the increasing weight assigned to the additional indicators leading BIP, which dominate the BIP component at larger forecast horizons. %In comparison, the effect of phase shifts—arising from the cyclical nature of the HP-filter and manifesting in the HP-C forecasts (shown in the left panels)—seems less systematic. 
\begin{figure}[H]\begin{center}\includegraphics[height=3in, width=4.5in]{./Figures/bk_h.pdf}\caption{Multivariate M-SSA: $h=0$ (nowcast, left) and $h=4$ (one year forecast, right).\label{bk_h}}\end{center}\end{figure}\begin{figure}[H]\begin{center}\includegraphics[height=3in, width=4.5in]{./Figures/multivar_vs_univar.pdf}\caption{Now- and forecasts of HP-BIP for $h=0,...,6$: HP-C (left) vs. M-SSA (right) across the financial crisis (top) and the pandemic (bottom).\label{multivar_vs_univar}}\end{center}\end{figure}

@@

There are various combinations of M-SSA components for tracking BIP. For illustration, we here select the single BIP component resulting from the M-SSA filter tracking HP-BIP. Table \eqref{p_val_hpbip_wc} reports HAC-adjusted p-values from regressions of the out-of-sample predictor on HP-BIP, indicating a strong link between the predictor and the target, out-of-sample. 
% latex table generated in R 4.2.2 by xtable 1.8-4 package
% Thu May  1 21:38:45 2025
\begin{table}[ht]
\centering
\begin{tabular}{rrrrrrrr}
  \hline
 & h=0 & h=1 & h=2 & h=3 & h=4 & h=5 & h=6 \\ 
  \hline
Shift=0 & 0.000 & 0.000 & 0.001 & 0.002 & 0.005 & 0.020 & 0.112 \\ 
  Shift=1 & 0.000 & 0.000 & 0.000 & 0.001 & 0.002 & 0.005 & 0.019 \\ 
  Shift=2 & 0.000 & 0.000 & 0.000 & 0.001 & 0.001 & 0.003 & 0.006 \\ 
  Shift=3 & 0.018 & 0.007 & 0.002 & 0.000 & 0.002 & 0.003 & 0.004 \\ 
  Shift=4 & 0.180 & 0.100 & 0.040 & 0.010 & 0.002 & 0.005 & 0.005 \\ 
  Shift=5 & 0.579 & 0.432 & 0.261 & 0.108 & 0.027 & 0.008 & 0.009 \\ 
   \hline
\end{tabular}
\caption{HAC adjusted p-values of regressions of BIP (M-SSA) component optimized for forecast horizons $h=0,...,6$  (the columns) on HP-BIP shifted forward by shift=0,...,5 (the rows). Out-of-sample span starting in Jan-2007, without pandemic.} 
\label{p_val_hpbip_wc}
\end{table}Our choice—focusing on the single M-SSA BIP component—emphasizes simplicity and interpretability, as future BIP and HP-BIP are connected through their shared low-frequency component. Additionally, the M-SSA BIP component is the most important predictor in the regression of the M-SSA outputs on BIP.  Finally, revisions due to quarterly updates of the regression equations are minimized by this predictor, and the estimates are more stable and robust over time\footnote{Other combinations can be experimented with using the M-SSA package. Regression weights of more complex designs, involving multiple components, tend to be more difficult to interpret due to the occurrence of negative weights and potential sign changes of the regressors over time, which may indicate overfitting or non-stationarity.} (see Section \eqref{revisions}). Note that information from the other  indicators is incorporated into the M-SSA BIP component through the multivariate filter design.\\

@@

% latex table generated in R 4.2.2 by xtable 1.8-4 package
% Thu May  1 21:38:45 2025
\begin{table}[ht]
\centering
\begin{tabular}{rrrrrrrr}
  \hline
 & h=0 & h=1 & h=2 & h=3 & h=4 & h=5 & h=6 \\ 
  \hline
Shift=0 & 0.084 & 0.047 & 0.017 & 0.005 & 0.002 & 0.001 & 0.038 \\ 
  Shift=1 & 0.230 & 0.160 & 0.094 & 0.032 & 0.005 & 0.001 & 0.004 \\ 
  Shift=2 & 0.801 & 0.666 & 0.414 & 0.173 & 0.026 & 0.005 & 0.006 \\ 
  Shift=3 & 0.876 & 0.961 & 0.917 & 0.683 & 0.248 & 0.068 & 0.039 \\ 
  Shift=4 & 0.370 & 0.819 & 0.990 & 0.953 & 0.533 & 0.117 & 0.030 \\ 
  Shift=5 & 0.353 & 0.802 & 0.983 & 0.943 & 0.571 & 0.239 & 0.043 \\ 
   \hline
\end{tabular}
\caption{HAC adjusted p-values of regressions of M-SSA component predictors optimized for forecast horizons $h=0,...,6$  (the columns) on BIP shifted forward by shift=0,...,5 (the rows). Out-of-sample span starting in Jan-2007, without pandemic.} 
\label{p_val_wc}
\end{table}% latex table generated in R 4.2.2 by xtable 1.8-4 package
% Thu May  1 21:38:45 2025
\begin{table}[ht]
\centering
\begin{tabular}{rrrrrrrr}
  \hline
 & h=0 & h=1 & h=2 & h=3 & h=4 & h=5 & h=6 \\ 
  \hline
Shift=0 & 0.978 & 0.942 & 0.893 & 0.855 & 0.860 & 0.906 & 0.962 \\ 
  Shift=1 & 1.082 & 1.060 & 1.005 & 0.926 & 0.877 & 0.880 & 0.914 \\ 
  Shift=2 & 1.083 & 1.090 & 1.077 & 1.019 & 0.952 & 0.918 & 0.919 \\ 
  Shift=3 & 1.034 & 1.059 & 1.081 & 1.063 & 1.010 & 0.970 & 0.954 \\ 
  Shift=4 & 0.995 & 1.029 & 1.069 & 1.077 & 1.035 & 0.986 & 0.953 \\ 
  Shift=5 & 0.993 & 1.030 & 1.075 & 1.093 & 1.062 & 1.011 & 0.967 \\ 
   \hline
\end{tabular}
\caption{rRMSEs of M-SSA components predictor benchmarked against the expanding mean of BIP. Out-of-sample span starting in Jan-2007, without pandemic.} 
\label{rRMSE_mSSA_comp_mean_wc}
\end{table}

@@

Revisions refer to changes in the historical values of a time series resulting from updates with new information. Revisions of the M-SSA component predictor can arise from data updates, which we do not analyze further here\footnote{Except for BIP, the other indicators are not or only slightly revised. Moreover, the smoothing effect of filters mitigates the effect of revisions when compared to direct forecasts.}, or from updates to predictor parameters. The latter can be further separated into changes in VAR parameters and regression weights. 
\subsection{Revisions: Regression Parameters}

@@

\subsection{Revisions: M-SSA}

@@

% latex table generated in R 4.2.2 by xtable 1.8-4 package
% Thu May  1 21:38:46 2025
\begin{table}[ht]
\centering
\begin{tabular}{rrrrrrrr}
  \hline
 & h=0 & h=1 & h=2 & h=3 & h=4 & h=5 & h=6 \\ 
  \hline
Shift=0 & 0.084 & 0.047 & 0.017 & 0.005 & 0.002 & 0.001 & 0.038 \\ 
  Shift=1 & 0.230 & 0.160 & 0.094 & 0.032 & 0.005 & 0.001 & 0.004 \\ 
  Shift=2 & 0.801 & 0.666 & 0.414 & 0.173 & 0.026 & 0.005 & 0.006 \\ 
  Shift=3 & 0.876 & 0.961 & 0.917 & 0.683 & 0.248 & 0.068 & 0.039 \\ 
  Shift=4 & 0.370 & 0.819 & 0.990 & 0.953 & 0.533 & 0.117 & 0.030 \\ 
  Shift=5 & 0.353 & 0.802 & 0.983 & 0.943 & 0.571 & 0.239 & 0.043 \\ 
   \hline
\end{tabular}
\caption{HAC adjusted p-values of regressions of M-SSA component predictors optimized for forecast horizons $h=0,...,6$  (the columns) on BIP shifted forward by shift=0,...,5 (the rows). Out-of-sample span starting in Jan-2007, without pandemic.} 
\label{p_val_wc5}
\end{table}% latex table generated in R 4.2.2 by xtable 1.8-4 package
% Thu May  1 21:38:46 2025
\begin{table}[ht]
\centering
\begin{tabular}{rrrrrrrr}
  \hline
 & h=0 & h=1 & h=2 & h=3 & h=4 & h=5 & h=6 \\ 
  \hline
Shift=0 & 0.978 & 0.942 & 0.893 & 0.855 & 0.860 & 0.906 & 0.962 \\ 
  Shift=1 & 1.082 & 1.060 & 1.005 & 0.926 & 0.877 & 0.880 & 0.914 \\ 
  Shift=2 & 1.083 & 1.090 & 1.077 & 1.019 & 0.952 & 0.918 & 0.919 \\ 
  Shift=3 & 1.034 & 1.059 & 1.081 & 1.063 & 1.010 & 0.970 & 0.954 \\ 
  Shift=4 & 0.995 & 1.029 & 1.069 & 1.077 & 1.035 & 0.986 & 0.953 \\ 
  Shift=5 & 0.993 & 1.030 & 1.075 & 1.093 & 1.062 & 1.011 & 0.967 \\ 
   \hline
\end{tabular}
\caption{rRMSEs of M-SSA components predictor benchmarked against the expanding mean of BIP. Out-of-sample span starting in Jan-2007, without pandemic.} 
\label{rRMSE_mSSA_comp_mean6}
\end{table}% latex table generated in R 4.2.2 by xtable 1.8-4 package
% Thu May  1 21:38:46 2025
\begin{table}[ht]
\centering
\begin{tabular}{rrrrrrrr}
  \hline
 & h=0 & h=1 & h=2 & h=3 & h=4 & h=5 & h=6 \\ 
  \hline
Shift=0 & 1.144 & 1.102 & 1.045 & 0.999 & 1.006 & 1.059 & 1.125 \\ 
  Shift=1 & 1.238 & 1.213 & 1.151 & 1.060 & 1.004 & 1.007 & 1.047 \\ 
  Shift=2 & 1.178 & 1.186 & 1.171 & 1.108 & 1.035 & 0.999 & 0.999 \\ 
  Shift=3 & 1.011 & 1.035 & 1.057 & 1.039 & 0.988 & 0.948 & 0.933 \\ 
  Shift=4 & 1.010 & 1.045 & 1.086 & 1.094 & 1.051 & 1.002 & 0.968 \\ 
  Shift=5 & 0.982 & 1.019 & 1.064 & 1.081 & 1.051 & 1.000 & 0.957 \\ 
   \hline
\end{tabular}
\caption{rRMSEs of M-SSA components predictor benchmarked against the direct forecasts. Out-of-sample span starting in Jan-2007, without pandemic.} 
\label{rRMSE_mSSA_comp_direct_without_covid7}
\end{table}% latex table generated in R 4.2.2 by xtable 1.8-4 package
% Thu May  1 21:38:46 2025
\begin{table}[ht]
\centering
\begin{tabular}{rrrrrrrr}
  \hline
 & h=0 & h=1 & h=2 & h=3 & h=4 & h=5 & h=6 \\ 
  \hline
Shift=0 & 0.855 & 0.855 & 0.855 & 0.855 & 0.855 & 0.855 & 0.855 \\ 
  Shift=1 & 0.874 & 0.874 & 0.874 & 0.874 & 0.874 & 0.874 & 0.874 \\ 
  Shift=2 & 0.919 & 0.919 & 0.919 & 0.919 & 0.919 & 0.919 & 0.919 \\ 
  Shift=3 & 1.023 & 1.023 & 1.023 & 1.023 & 1.023 & 1.023 & 1.023 \\ 
  Shift=4 & 0.985 & 0.985 & 0.985 & 0.985 & 0.985 & 0.985 & 0.985 \\ 
  Shift=5 & 1.010 & 1.010 & 1.010 & 1.010 & 1.010 & 1.010 & 1.010 \\ 
   \hline
\end{tabular}
\caption{rRMSEs direct forecasts benchmarked against the expanding mean. Out-of-sample span starting in Jan-2007, without pandemic.} 
\label{rRMSE_mSSA_direct_mean_without_covid8}
\end{table}

>>>>

Extending the `classic' direct forecast design, we here consider filtered indicators, based on the one-sided HP(160), denoted as HP-C: the filtered indicators are explanatory variables for the regressions on forward-shifted BIP\footnote{Katja and Simon: for each forecast horizon $h=0,1,2,3$ I'm using corresponding forecasts of the real-time filter output, assuming the data to be white noise (assuming an ARMA model does not lead to significantly different results). I think we might ignore these technical details in the paper.}. As in the previous section, we rely on all indicators as well as the full sample estimate for illustration, see Figs.\eqref{direct_hp_forecasts} (all observations with exclusion of the pandemic) and \eqref{direct_hp_forecasts_financial_crisis} (financial crisis). A comparison with the classic direct forecasts in Fig.\eqref{direct_wc_financial_crisis} indicates that the new predictors are slightly less retarded (`faster'). 
\begin{figure}[H]\begin{center}\includegraphics[height=3in, width=4.5in]{./Figures/direct_hp_forecasts.pdf}\caption{Direct HP forecasts: entire data set with exclusion of the pandemic.\label{direct_hp_forecasts}}\end{center}\end{figure}\begin{figure}[H]\begin{center}\includegraphics[height=3in, width=4.5in]{./Figures/direct_hp_forecasts_financial_crisis.pdf}\caption{Direct HP forecasts over the curse of the financial crisis.\label{direct_hp_forecasts_financial_crisis}}\end{center}\end{figure}Moreover, the F-statistics are larger than for the direct forecasts, see Table \eqref{f_stat}, but the predictor remains statistically insignificant at forecast horizons exceeding two quarters (even though the evaluation is conducted in-sample and based on all available indicators). 
% latex table generated in R 4.2.2 by xtable 1.8-4 package
% Wed May  7 07:36:27 2025
\begin{table}[ht]
\centering
\begin{tabular}{rrrrr}
  \hline
 & h=0 & h=1 & h=2 & h=3 \\ 
  \hline
Direct forecast & 7.188 & 5.285 & 3.574 & 0.460 \\ 
  Direct HP forecast & 21.886 & 8.415 & 4.810 & 1.834 \\ 
   \hline
\end{tabular}
\caption{F-statistics of direct forecasts and direct HP forecasts for forecast horizons $h=0,1,2,3$: fully in-sample.  } 
\label{f_stat}
\end{table}We now propose a more refined causal filter design aiming for statistical significance of the resulting predictor at longer forecast horizons, out-of-sample.

@@

% Wed May  7 07:36:27 2025

@@

Target correlation & 0.55 & 0.68 & 0.64 & 0.69 \\ 
  HT & 8.67 & 6.50 & 7.43 & 3.71 \\ 

@@

M-SSA predictors are compared to the traditional univariate HP-C in Figs. \eqref{multivar_vs_univar} and \eqref{mssa_hpc_financial_crisis}. In the first figure, we compare both filters across the periods of the financial crisis (top panels) and the COVID-19 pandemic (bottom panels), for forecast horizons $h=0,...,6$. The nowcast from the previous section is included for reference. In the second figure we compare both filters to the forward-shifted BIP for shifts larger than two quarters. All series are standardized to facilitate visual comparison. The primary distinction of the M-SSA approach is the progressive left-shift of the predictor as a function of $h$; this shift becomes more pronounced with increasing $h$ and appears to be consistent across all levels, including the peaks and troughs of the series. Conversely, the positions of peaks and troughs in the HP-C series appear to remain largely unaffected by variations in $h$. The more pronounced and systematic left-shift observed in the M-SSA forecasts can be partly attributed to the increasing weight assigned to the additional indicators leading BIP, which dominate the BIP component at larger forecast horizons. %In comparison, the effect of phase shifts—arising from the cyclical nature of the HP-filter and manifesting in the HP-C forecasts (shown in the left panels)—seems less systematic. 
\begin{figure}[H]\begin{center}\includegraphics[height=3in, width=4.5in]{./Figures/bk_h.pdf}\caption{Multivariate M-SSA: $h=0$ (nowcast, left) and $h=4$ (one year forecast, right).\label{bk_h}}\end{center}\end{figure}\begin{figure}[H]\begin{center}\includegraphics[height=3in, width=4.5in]{./Figures/multivar_vs_univar.pdf}\caption{Now- and forecasts of HP-BIP for $h=0,...,6$: HP-C (top) vs. M-SSA (bottom) across the financial crisis (left) and the pandemic (right).\label{multivar_vs_univar}}\end{center}\end{figure}
\begin{figure}[H]\begin{center}\includegraphics[height=3in, width=4.5in]{./Figures/mssa_hpc_financial_crisis.pdf}\caption{Forward-shifted BIP (black), M-SSA (blue) and HP-C (red) for $h=3,4,5,6$ across the financial crisis.\label{mssa_hpc_financial_crisis}}\end{center}\end{figure}

@@

There are various combinations of M-SSA components for tracking BIP. For illustration, we here select the single BIP component resulting from the M-SSA filter tracking HP-BIP. Table \eqref{p_val_hpbip_wc} reports HAC-adjusted p-values from regressions of the out-of-sample predictor on HP-BIP, indicating a strong link between the predictor and the target, out-of-sample. 
% latex table generated in R 4.2.2 by xtable 1.8-4 package
% Wed May  7 07:36:30 2025
\begin{table}[ht]
\centering
\begin{tabular}{rrrrrrrr}
  \hline
 & h=0 & h=1 & h=2 & h=3 & h=4 & h=5 & h=6 \\ 
  \hline
Shift=0 & 0.000 & 0.000 & 0.001 & 0.002 & 0.005 & 0.020 & 0.112 \\ 
  Shift=1 & 0.000 & 0.000 & 0.000 & 0.001 & 0.002 & 0.005 & 0.019 \\ 
  Shift=2 & 0.000 & 0.000 & 0.000 & 0.001 & 0.001 & 0.003 & 0.006 \\ 
  Shift=3 & 0.018 & 0.007 & 0.002 & 0.000 & 0.002 & 0.003 & 0.004 \\ 
  Shift=4 & 0.180 & 0.100 & 0.040 & 0.010 & 0.002 & 0.005 & 0.005 \\ 
  Shift=5 & 0.579 & 0.432 & 0.261 & 0.108 & 0.027 & 0.008 & 0.009 \\ 
   \hline
\end{tabular}
\caption{HAC adjusted p-values of regressions of BIP (M-SSA) component optimized for forecast horizons $h=0,...,6$  (the columns) on HP-BIP shifted forward by shift=0,...,5 (the rows). Out-of-sample span starting in Jan-2007, without pandemic.} 
\label{p_val_hpbip_wc}
\end{table}Our choice—focusing on the single M-SSA BIP component—emphasizes simplicity and interpretability, as future BIP and HP-BIP are connected through their shared low-frequency component. Additionally, the M-SSA BIP component is the most important predictor in the regression of the M-SSA outputs on BIP.  Finally, revisions due to quarterly updates of the regression equations are minimized by this predictor, and the estimates are more stable and robust over time\footnote{Other combinations can be experimented with using the M-SSA package. Regression weights of more complex designs, involving multiple components, tend to be more difficult to interpret due to the occurrence of negative weights and potential sign changes of the regressors over time, which may indicate overfitting or non-stationarity.} (see Section \eqref{revisions}). Note that information from the other  indicators is incorporated into the M-SSA BIP component through the multivariate filter design.\\

@@

% latex table generated in R 4.2.2 by xtable 1.8-4 package
% Wed May  7 07:36:30 2025
\begin{table}[ht]
\centering
\begin{tabular}{rrrrrrrr}
  \hline
 & h=0 & h=1 & h=2 & h=3 & h=4 & h=5 & h=6 \\ 
  \hline
Shift=0 & 0.084 & 0.047 & 0.017 & 0.005 & 0.002 & 0.001 & 0.038 \\ 
  Shift=1 & 0.230 & 0.160 & 0.094 & 0.032 & 0.005 & 0.001 & 0.004 \\ 
  Shift=2 & 0.801 & 0.666 & 0.414 & 0.173 & 0.026 & 0.005 & 0.006 \\ 
  Shift=3 & 0.876 & 0.961 & 0.917 & 0.683 & 0.248 & 0.068 & 0.039 \\ 
  Shift=4 & 0.370 & 0.819 & 0.990 & 0.953 & 0.533 & 0.117 & 0.030 \\ 
  Shift=5 & 0.353 & 0.802 & 0.983 & 0.943 & 0.571 & 0.239 & 0.043 \\ 
   \hline
\end{tabular}
\caption{HAC adjusted p-values of regressions of M-SSA component predictors optimized for forecast horizons $h=0,...,6$  (the columns) on BIP shifted forward by shift=0,...,5 (the rows). Out-of-sample span starting in Jan-2007, without pandemic.} 
\label{p_val_wc}
\end{table}% latex table generated in R 4.2.2 by xtable 1.8-4 package
% Wed May  7 07:36:30 2025
\begin{table}[ht]
\centering
\begin{tabular}{rrrrrrrr}
  \hline
 & h=0 & h=1 & h=2 & h=3 & h=4 & h=5 & h=6 \\ 
  \hline
Shift=0 & 0.978 & 0.942 & 0.893 & 0.855 & 0.860 & 0.906 & 0.962 \\ 
  Shift=1 & 1.082 & 1.060 & 1.005 & 0.926 & 0.877 & 0.880 & 0.914 \\ 
  Shift=2 & 1.083 & 1.090 & 1.077 & 1.019 & 0.952 & 0.918 & 0.919 \\ 
  Shift=3 & 1.034 & 1.059 & 1.081 & 1.063 & 1.010 & 0.970 & 0.954 \\ 
  Shift=4 & 0.995 & 1.029 & 1.069 & 1.077 & 1.035 & 0.986 & 0.953 \\ 
  Shift=5 & 0.993 & 1.030 & 1.075 & 1.093 & 1.062 & 1.011 & 0.967 \\ 
   \hline
\end{tabular}
\caption{rRMSEs of M-SSA components predictor benchmarked against the expanding mean of BIP. Out-of-sample span starting in Jan-2007, without pandemic.} 
\label{rRMSE_mSSA_comp_mean_wc}
\end{table}

@@

Revisions refer to changes in the historical values of a time series resulting from updates with new information. Revisions of the M-SSA component predictor can arise from data updates, which we do not analyze further here\footnote{Except for BIP, the other indicators are not or only slightly revised. Moreover, the smoothing effect of filters mitigates the effect of revisions when compared to direct forecasts.}, or from updates to predictor parameters. The latter can be further separated into changes in VAR parameters and regression weights.

\subsection{Revisions: Regression Equation Only}

@@

\subsection{Revisions: M-SSA Only}

@@

% latex table generated in R 4.2.2 by xtable 1.8-4 package
% Wed May  7 07:36:32 2025
\begin{table}[ht]
\centering
\begin{tabular}{rrrrrrrr}
  \hline
 & h=0 & h=1 & h=2 & h=3 & h=4 & h=5 & h=6 \\ 
  \hline
Shift=0 & 0.084 & 0.047 & 0.017 & 0.005 & 0.002 & 0.001 & 0.038 \\ 
  Shift=1 & 0.230 & 0.160 & 0.094 & 0.032 & 0.005 & 0.001 & 0.004 \\ 
  Shift=2 & 0.801 & 0.666 & 0.414 & 0.173 & 0.026 & 0.005 & 0.006 \\ 
  Shift=3 & 0.876 & 0.961 & 0.917 & 0.683 & 0.248 & 0.068 & 0.039 \\ 
  Shift=4 & 0.370 & 0.819 & 0.990 & 0.953 & 0.533 & 0.117 & 0.030 \\ 
  Shift=5 & 0.353 & 0.802 & 0.983 & 0.943 & 0.571 & 0.239 & 0.043 \\ 
   \hline
\end{tabular}
\caption{HAC adjusted p-values of regressions of M-SSA component predictors optimized for forecast horizons $h=0,...,6$  (the columns) on BIP shifted forward by shift=0,...,5 (the rows). Out-of-sample span starting in Jan-2007, without pandemic.} 
\label{p_val_wc5}
\end{table}% latex table generated in R 4.2.2 by xtable 1.8-4 package
% Wed May  7 07:36:32 2025
\begin{table}[ht]
\centering
\begin{tabular}{rrrrrrrr}
  \hline
 & h=0 & h=1 & h=2 & h=3 & h=4 & h=5 & h=6 \\ 
  \hline
Shift=0 & 0.978 & 0.942 & 0.893 & 0.855 & 0.860 & 0.906 & 0.962 \\ 
  Shift=1 & 1.082 & 1.060 & 1.005 & 0.926 & 0.877 & 0.880 & 0.914 \\ 
  Shift=2 & 1.083 & 1.090 & 1.077 & 1.019 & 0.952 & 0.918 & 0.919 \\ 
  Shift=3 & 1.034 & 1.059 & 1.081 & 1.063 & 1.010 & 0.970 & 0.954 \\ 
  Shift=4 & 0.995 & 1.029 & 1.069 & 1.077 & 1.035 & 0.986 & 0.953 \\ 
  Shift=5 & 0.993 & 1.030 & 1.075 & 1.093 & 1.062 & 1.011 & 0.967 \\ 
   \hline
\end{tabular}
\caption{rRMSEs of M-SSA components predictor benchmarked against the expanding mean of BIP. Out-of-sample span starting in Jan-2007, without pandemic.} 
\label{rRMSE_mSSA_comp_mean6}
\end{table}% latex table generated in R 4.2.2 by xtable 1.8-4 package
% Wed May  7 07:36:32 2025
\begin{table}[ht]
\centering
\begin{tabular}{rrrrrrrr}
  \hline
 & h=0 & h=1 & h=2 & h=3 & h=4 & h=5 & h=6 \\ 
  \hline
Shift=0 & 1.144 & 1.102 & 1.045 & 0.999 & 1.006 & 1.059 & 1.125 \\ 
  Shift=1 & 1.238 & 1.213 & 1.151 & 1.060 & 1.004 & 1.007 & 1.047 \\ 
  Shift=2 & 1.178 & 1.186 & 1.171 & 1.108 & 1.035 & 0.999 & 0.999 \\ 
  Shift=3 & 1.011 & 1.035 & 1.057 & 1.039 & 0.988 & 0.948 & 0.933 \\ 
  Shift=4 & 1.010 & 1.045 & 1.086 & 1.094 & 1.051 & 1.002 & 0.968 \\ 
  Shift=5 & 0.982 & 1.019 & 1.064 & 1.081 & 1.051 & 1.000 & 0.957 \\ 
   \hline
\end{tabular}
\caption{rRMSEs of M-SSA components predictor benchmarked against the direct forecasts. Out-of-sample span starting in Jan-2007, without pandemic.} 
\label{rRMSE_mSSA_comp_direct_without_covid7}
\end{table}% latex table generated in R 4.2.2 by xtable 1.8-4 package
% Wed May  7 07:36:32 2025
\begin{table}[ht]
\centering
\begin{tabular}{rrrrrrrr}
  \hline
 & h=0 & h=1 & h=2 & h=3 & h=4 & h=5 & h=6 \\ 
  \hline
Shift=0 & 0.855 & 0.855 & 0.855 & 0.855 & 0.855 & 0.855 & 0.855 \\ 
  Shift=1 & 0.874 & 0.874 & 0.874 & 0.874 & 0.874 & 0.874 & 0.874 \\ 
  Shift=2 & 0.919 & 0.919 & 0.919 & 0.919 & 0.919 & 0.919 & 0.919 \\ 
  Shift=3 & 1.023 & 1.023 & 1.023 & 1.023 & 1.023 & 1.023 & 1.023 \\ 
  Shift=4 & 0.985 & 0.985 & 0.985 & 0.985 & 0.985 & 0.985 & 0.985 \\ 
  Shift=5 & 1.010 & 1.010 & 1.010 & 1.010 & 1.010 & 1.010 & 1.010 \\ 
   \hline
\end{tabular}
\caption{rRMSEs direct forecasts benchmarked against the expanding mean. Out-of-sample span starting in Jan-2007, without pandemic.} 
\label{rRMSE_mSSA_direct_mean_without_covid8}
\end{table}
